---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First, we are gonna set our constants and check if the working directory exists. Then we'll set our working environment.

```{r setwd, echo = TRUE}
#Here, you can change the working directory
MAIN_DIRECTORY <- "C:/Project1ReproducibleResearch";
DATA_URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip";
DATA_ZIP_FILE <- "activity.zip";
DATA_UNZIP_DIRECTORY <- "activity_data";
DATA_FILE <- "activity.csv";

#Verifying if main directory exists.
if(!file.exists(MAIN_DIRECTORY)){ dir.create(MAIN_DIRECTORY); }
setwd(MAIN_DIRECTORY);
```

Then we'll download the data and then load the data into dataframe.

```{r load, echo = TRUE}
#Verifying if data file exists.
if(!file.exists(DATA_ZIP_FILE)){
  download.file(DATA_URL, DATA_ZIP_FILE);
}
#Verifying if directory exists.
if(!file.exists(DATA_UNZIP_DIRECTORY)){
  unzip(DATA_ZIP_FILE, exdir = DATA_UNZIP_DIRECTORY);
}

#Read csv data
activity_data <- read.csv(file.path(DATA_UNZIP_DIRECTORY, DATA_FILE));

#Cast the date attribute
activity_data$date <- as.Date(activity_data$date);
```

Let's show how looks like the data
```{r show, echo = TRUE}
head(activity_data);
```

## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
